---
title: "Assignment1"
subtitle: "STAE04/DABN19: Data Visualization"
author: "Roshan Talimi"
date: "`r Sys.Date()`"
papersize: a4
geometry: margin=4cm
colorlinks: true
output:
  pdf_document:
    number_sections: true
---

```{r include = FALSE}
#https://raw.githubusercontent.com/stat-lu/dataviz/main/template.Rmd"
```

```{r setup, include = FALSE}
# Setup options for R Markdown
knitr::opts_chunk$set(
  echo       = FALSE,    # Do not print code
  warning    = FALSE,    # Suppress warnings
  message    = FALSE,    # Suppress messages
  fig.align  = "center", # Center figures
  fig.width  = 2.7,      # Good standard figure width for single-panel figures
  fig.height = 2.4       # Good standard figure height for single-panel figures
)

library(tidyverse)

# Set a theme for ggplot2
theme_set(theme_light(base_size = 15))

# Set options
options(
  digits = 3, # limit the number of significant digits
  width  = 63 # limit the width of code output
)
```

```{r echo=FALSE, eval=FALSE}
head(starwars)
```

Task 1

```{r echo=TRUE}
starwars_tatooine_summary <-
  starwars %>% 
  filter(homeworld == "Tatooine") %>% 
  drop_na(any_of("mass")) %>% 
  group_by(eye_color) %>% 
  summarize(avg_mass = mean(mass))
```


```{r, fig.cap = "Average weight of Tatooine born Star Wars charachters by eye color."}
ggplot(starwars_tatooine_summary, aes(eye_color, avg_mass)) +
  geom_col() +
  xlab("Eye Color") +
  ylab("Average Weight")
```


Task 2

Cases of tuberculosis in three countries across two years.

```{r echo=FALSE}
knitr::kable(table4a)
```

The above dataset is not tidy because:

1. It does not have a row for every observation(e.g. Afghanistan 1999 745)
2. It does not have a column for every variable (e.g. Year or Cases)

Our 3 variables in this data are Country, Year and Tuberculosis Cases.

```{r echo=TRUE}
table4a_tidy <-
  table4a %>%
  pivot_longer(
    c("1999", "2000"), 
    names_to = "year", 
    values_to = "cases"
  )
```

The tidy version of the data looks like this:

```{r echo=FALSE}
knitr::kable(table4a_tidy)
```
Task 3

Population for the countries during that period is:

```{r}
table4b_tidy <-
  table4b %>%
  pivot_longer(
    c("1999", "2000"), 
    names_to = "year", 
    values_to = "population"
  )
```

```{r echo=FALSE}
knitr::kable(table4b_tidy)
```

We can now calculate the rate of tuberculosis cases for population by country and year

```{r echo=FALSE}
table4 <- left_join(table4b_tidy, table4a_tidy)
table4 %>% mutate(rate = cases/population)
```






