---
title: "The Class of Fall 2021"
subtitle: "STAE04: Data Visualization"
author: "Johan Larsson"
affiliation: "Department of Statistics, Lund University"
date: "`r Sys.Date()`"
toc: false
papersize: a4
geometry: margin=4cm
colorlinks: true
output:
  pdf_document:
    number_sections: true
---

```{r setup, include = FALSE}
# This is the setup chunk. It can be used to setup general options and load
# packages that you make use of during the report.
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 2.5,
  fig.height = 2.2
)

library(tidyverse)

# set ggplot2 theme options
theme_set(theme_grey(base_size = 10))

options(
  digits = 3, # limit number of significant digits
  width = 63  # limit width of code
)
```

# Introduction

```{r, echo = FALSE}
library(tidyverse)

tmp <- read_csv("stae21fall.csv")

dat <-
  read_csv("stae21fall.csv") %>%
  select(c(6, 8, 10, 12, 14)) %>%
  rename_with(~ c(
    "Occupation",
    "Workload",
    "Dataviz_Experience",
    "R_Experience",
    "Learnt_About_Course"
  )) %>%
  mutate(
    Dataviz_Experience =
      fct_relabel(
        Dataviz_Experience,
        ~ str_extract(., "^[[:alnum:]]+")
      ),
    R_Experience =
      fct_relabel(
        R_Experience,
        ~ str_extract(., "^[[:alnum:]]+")
      ),
    Dataviz_Experience = factor(
      Dataviz_Experience,
      c("novice", "beginner", "intermediate"),
      ordered = TRUE
    ),
    R_Experience = factor(
      R_Experience,
      c("novice", "beginner", "intermediate"),
      ordered = TRUE
    ),
    Learnt_About_Course = factor(
      Learnt_About_Course,
      labels = c(
        "antagning.se",
        "relative, friend, or colleague",
        "Social Media",
        "Dept.Website"
      )
    )
  )
```

Let's take a look at the survey results. `r nrow(dat)` of you have completed
the survey to date. I have pulled your data from Canvas and formatted it to make
it compatible with ggplot2. 

# Occupation

Let's start by looking at what you do. In \autoref{fig:occupation}, we see 
that there is a fairly even split between full-time students and employees in
the group. Some of you responded "other", which presumably indicates that you
are either working part-time or studying part-time.

```{r occupation, fig.cap = "The occupation of the students in the class.", fig.width = 5}
ggplot(dat, aes(Occupation)) +
  geom_bar() +
  labs(y = "Count") +
  scale_x_discrete(
    labels = function(x) sub(",", ",\n", x, fixed = TRUE))
```

# Planned Hours of Work

By looking at \autoref{fig:workload}, it appears that there is quite a lot of
variability in how much time you plan to spend on the course each week. The mode
of the distribution (most common value), however, is ten, which is spot on in
terms of what the intended workload of the course is. It is perhaps, however, a
little worrying that many of you plan to spend only five hours each week on the
course, since this means you'll have to be twice as efficient as my estimate!

```{r workload, fig.cap = "The number of hours of work you plan to work on the course each week.", fig.width = 5, fig.height = 2}
ggplot(dat, aes(Workload, fill = Occupation)) +
  geom_dotplot(
    binwidth = 1,
    binpositions = "all",
    stackgroups = TRUE
  ) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "right"
  ) +
  xlab("hours") +
  scale_fill_brewer(type = "qual")
```

# Experience with Data Visualization and R

As expected, the level of experience with R and with Data Visualization appears
to be correlated (\autoref{fig:experience}). It is perhaps also noteworthy that
none of you consider yourselves to be experts in either data visualization or R,
which to me says that you're likely a good match for the course. The large
majority of the group also place themselves in the category *beginner* in terms
of experience with data visualization. 

```{r experience, fig.width = 4.5, fig.height = 3.5, fig.cap = "Experience with R and data visualization. Notice that the overlapping labels are not quite ideal here. (I did not have time to fix it properly, sorry!)"}
library(ggmosaic)

dat %>%
  drop_na(R_Experience, Dataviz_Experience) %>%
  ggplot() +
  geom_mosaic(aes(
    product(R_Experience, Dataviz_Experience),
    fill = R_Experience
  )) +
  theme(legend.position = "none") +
  labs(y = "R Experience", x = "Dataviz Experience")
```

# Expectations

Many of you of course have expectations related to visualizations, both in terms
of practically constructing visualizations and also how to do so effectively,
for instance by telling stories using visualizations. These are both among the
primary goals of this course, so I hope you'll find that we satisfy these
expectations adequately.

Some of you also mentioned specific skills or types of visualizations, such as
animated visualizations and maps, which are both part of the course. In
particular, we will focus quite a bit on maps in the course. Please note that
even if we only cover the basics on some of these topics, you will find that the
skills you learn are easy to extend to new types of visualizations.

Some of you also wrote specifically about learning to program in R. To be fair,
this is not a course that's dedicated to programming in R. We are using R (and
ggplot2) because they are fantastic tools for visualizing data, but will only
cover the bare necessities for doing visualization work in R. Many aspects that
would be part of an introductory course to R, such as loop constructs, will not
be covered at all, so please do not expect to come out expert programmers after
you've finished here!

Another thing that came up when I read your expectations is building statistical
models. We will only talk about statistical modeling in this course at a very
basic level and only consider topics that you are expected to have learned in
your introductory courses to statistics. We will cover visualizing estimates and
predictions from linear regression, but that's about it.

Good luck on the rest of the course!
