---
title: "Assignment 3"
subtitle: "STAE04/DABN19: Data Visualization"
author: "Roshan Talimi"
date: "`r Sys.Date()`"
papersize: a4
geometry: margin=4cm
colorlinks: true
output:
  pdf_document:
    number_sections: true
---

```{r include = FALSE}
#https://raw.githubusercontent.com/stat-lu/dataviz/main/template.Rmd"
```

```{r setup, include = FALSE}
# Setup options for R Markdown
knitr::opts_chunk$set(
  echo       = FALSE,    # Do not print code
  warning    = FALSE,    # Suppress warnings
  message    = FALSE,    # Suppress messages
  fig.align  = "center", # Center figures
  fig.width  = 2.7,      # Good standard figure width for single-panel figures
  fig.height = 2.4       # Good standard figure height for single-panel figures
)

library(tidyverse)
library(Ecdat)

# Set a theme for ggplot2
theme_set(theme_light(base_size = 15))

# Set options
options(
  digits = 3, # limit the number of significant digits
  width  = 63 # limit the width of code output
)
```

Task 1: Getting a Feel for the Data

The wages data set consists of 4165 observations/individuals and 12 variables.
The source of the data set is a panel of 595 observations from 1976 to 1982 in United States.

The variables are:

* exp:        years of full-time work experience 
* wks:        weeks worked
* bluecol:    blue collar 
* ind:        works in a manufacturing industry
* south:      resides in the south
* smsa:       resides in a standard metropolitan statistical are
* married:    married
* sex:        a factor with levels (male,female)
* union:      individual's wage set by a union contract
* ed:         years of education
* black:      is the individual black
* lwage:      logarithm of wage ($)


Source:
Cornwell, C. and P. Rupert (1988) “Efficient estimation with panel data: an empirical comparison of instrumental variables estimators”, Journal of Applied Econometrics, 3, 149–155.

Panel study of income dynamics.

References:
Baltagi, Badi H. (2003) Econometric analysis of panel data, John Wiley and sons, https://www.wiley.com/legacy/wileychi/baltagi/.


```{r echo=FALSE}
data("Wages")
Wages <- Wages %>% mutate(wage = exp(lwage))
```


```{r echo=TRUE}

```


Before you begin, add a new variable to the data set containing the wages at its original (non-transformed) scale using mutate(). Note that the lwage column in the data set represents wages logarithmized with the natural logarithm. Name the new variable wage.

Hint: What is the inverse of the (natural) logarithm? e^x

```{r echo=FALSE}
Wages

```

```{r, fig.cap = "Years of experience and wage"}
Wages %>% ggplot(aes(exp, wage)) + geom_point()  #add jidder
Wages %>% ggplot(aes(exp, lwage)) + geom_point() #add jidder
Wages %>% ggplot(aes(wage, lwage)) + geom_point()
```


Task 2

Task 2: Weeks Worked and Wages
Let’s move on to take a look at wages and its relationship to the number of weeks worked in a year. 
Instead of looking at the entire data set, however, we will only consider individuals working for 40 or more weeks. 
Filter out the remaining observations and then construct your plot.

Comment on the pattern in your plot—what shape is it? Speculate as to why the pattern looks the way it does.

```{r echo=TRUE}
Wages %>% filter(wks >= 40) %>% ggplot(aes(exp, wage)) + geom_point()
Wages 
```

```{r echo=FALSE}

```


Task 3

```{r}
Wages <-
  Wages %>%
  mutate(wage_levels = cut(
    wage,
    c(0, 750, 1500, max(wage)),
    labels = c("Low", "Mid", "High")
  ))
```


Now, your job is to construct a plot of the relationship between wage levels (wage_levels) and one of the following variables:

whether the job is a blue-collar job or not,
whether the job is in a manufacturing industry,
whether the person resides in the south,
whether the person is married,
the sex of the individual,
whether the wage is set by a union contract, and
whether the individual is black or not.
Explore the various relationships using different visualization techniques and choose one that you find to be particularly interesting to include in your final report.


```{r echo=FALSE}
Wages %>% filter(wks >= 40) %>% ggplot(aes(exp, wage, color = wage_levels)) + geom_point()
```



```{r echo=FALSE}

```






