---
title: "Assignment 3"
subtitle: "STAE04/DABN19: Data Visualization"
author: "Roshan Talimi"
date: "`r Sys.Date()`"
papersize: a4
geometry: margin=4cm
colorlinks: true
output:
  pdf_document:
    number_sections: true
---

```{r include = FALSE}
library(tidyverse)
library(Ecdat)
theme_set(theme_light(base_size = 15))
#https://raw.githubusercontent.com/stat-lu/dataviz/main/template.Rmd"
```

# Task 1: Getting a Feel for the Data

The wages data set consists of 4165 observations/individuals and 12 variables.
The source of the data set is a panel of 595 observations from 1976 to 1982 in United States.

The variables are:

* exp:        years of full-time work experience 
* wks:        weeks worked
* bluecol:    blue collar 
* ind:        works in a manufacturing industry
* south:      resides in the south
* smsa:       resides in a standard metropolitan statistical are
* married:    married
* sex:        a factor with levels (male,female)
* union:      individual's wage set by a union contract
* ed:         years of education
* black:      is the individual black
* lwage:      logarithm of wage ($)



Source:
Cornwell, C. and P. Rupert (1988) “Efficient estimation with panel data: an empirical comparison of instrumental variables estimators”, Journal of Applied Econometrics, 3, 149–155.

Panel study of income dynamics.

References:
Baltagi, Badi H. (2003) Econometric analysis of panel data, John Wiley and sons, https://www.wiley.com/legacy/wileychi/baltagi


```{r echo=FALSE, include = FALSE}
data("Wages")
Wages <- Wages %>% mutate(wage = exp(lwage))
```

I'll add a new variable called wage that is the exponent of lwage so the real wage.
What is the inverse of the (natural) logarithm? e^ln(x) = x



Relationship between years of experience and wage can be seen below:

```{r echo=FALSE, error=FALSE, message=FALSE}
Wages %>% 
  ggplot(aes(exp, wage)) + 
  geom_jitter() + 
  geom_smooth(method = "lm") + 
  labs(title = "Wage vs years of experience",
       x = "Years of full-time work experience ",
       y = "Wage")
```

The correlation between wage and years of experience is 0.2 so not a very significant correlation.

```{r echo=FALSE, include = FALSE}
cor(Wages$exp, Wages$wage)
```





# Task 2: Weeks Worked and Wages

Let’s look at wages and its relationship to the number of weeks worked in a year. 
Instead of looking at the entire data set, we will only consider individuals working for 40 or more weeks. 
They make up around 91% of the data. 


```{r echo=FALSE, error=FALSE, message=FALSE}
Wages %>% 
  filter(wks >= 40) %>% 
  ggplot(aes(exp, lwage)) + 
  geom_jitter() + 
  geom_smooth(method = "lm") +
  labs(title = "Wage vs years of experience for people with over 40 weeks worked",
       x = "Weeks worked",
       y = "Natural Logarith of Wage")
```

I don't see a lot of change for people working full time or less.
The correlations doesn't change much for this group in particular.

# Task 3

```{r echo=FALSE, include = FALSE}
Wages <-
  Wages %>%
  mutate(wage_levels = cut(
    wage,
    c(0, 750, 1500, max(wage)),
    labels = c("Low", "Mid", "High")
  ))
```


The relationship between wage levels and the sex of the individuals:

```{r echo=FALSE, error=FALSE, message=FALSE}
Wages %>% 
  ggplot(aes(exp, wage, color = wage_levels)) + 
  geom_jitter() + 
  facet_wrap(~sex, nrow = 2) + 
  labs(title = "Wage vs years of experience for male and females",
       x = "Years of full-time work experience ",
       y = "Wage",
       colour = "Gender")
```
We see clearly that during 1976 to 1982 there is a distinction between males and females.




