---
title: "Assignment 3"
subtitle: "STAE04/DABN19: Data Visualization"
author: "Roshan Talimi"
date: "`r Sys.Date()`"
papersize: a4
geometry: margin=4cm
colorlinks: true
output:
  pdf_document:
    number_sections: true
---

```{r include = FALSE}
#https://raw.githubusercontent.com/stat-lu/dataviz/main/template.Rmd"
```

```{r setup, include = FALSE}
# Setup options for R Markdown
knitr::opts_chunk$set(
  echo       = FALSE,    # Do not print code
  warning    = FALSE,    # Suppress warnings
  message    = FALSE,    # Suppress messages
  fig.align  = "center", # Center figures
  fig.width  = 2.7,      # Good standard figure width for single-panel figures
  fig.height = 2.4       # Good standard figure height for single-panel figures
)

library(tidyverse)
library(Ecdat)

# Set a theme for ggplot2
theme_set(theme_light(base_size = 15))

# Set options
options(
  digits = 3, # limit the number of significant digits
  width  = 63 # limit the width of code output
)
```

Task 1: Getting a Feel for the Data

The wages data set consists of 4165 observations/individuals and 12 variables.
The source of the data set is a panel of 595 observations from 1976 to 1982 in United States.

The variables are:

* exp:        years of full-time work experience 
* wks:        weeks worked
* bluecol:    blue collar 
* ind:        works in a manufacturing industry
* south:      resides in the south
* smsa:       resides in a standard metropolitan statistical are
* married:    married
* sex:        a factor with levels (male,female)
* union:      individual's wage set by a union contract
* ed:         years of education
* black:      is the individual black
* lwage:      logarithm of wage ($)

Source:
Cornwell, C. and P. Rupert (1988) “Efficient estimation with panel data: an empirical comparison of instrumental variables estimators”, Journal of Applied Econometrics, 3, 149–155.

Panel study of income dynamics.

References:
Baltagi, Badi H. (2003) Econometric analysis of panel data, John Wiley and sons, https://www.wiley.com/legacy/wileychi/baltagi/.


```{r echo=FALSE}
data("Wages")
Wages <- Wages %>% mutate(wage = exp(lwage))
```

I'll add a new variable called wage that is the exponent of lwage so the real wage.
What is the inverse of the (natural) logarithm? e^ln(x) = x0

Relationship between years of experience and wage can be seen below:

```{r echo=TRUE}
Wages %>% ggplot(aes(exp, wage)) + geom_jitter() + geom_smooth(method = "lm")
Wages %>% ggplot(aes(exp, lwage)) + geom_jitter() + geom_smooth(method = "lm")
```

The correlation between wage and years of experience is 0.2 so not a very significant correlation.

```{r echo=FALSE}
cor(Wages$exp, Wages$wage)
```

```{r, fig.cap = "Years of experience and wage"}
Wages %>% ggplot(aes(exp, lwage)) + geom_jitter() + geom_smooth(method = "lm")
```

```{r echo=FALSE}
cor(Wages$exp, Wages$lwage)
```

Task 2: Weeks Worked and Wages

Let’s look at wages and its relationship to the number of weeks worked in a year. 
Instead of looking at the entire data set, however, we will only consider individuals working for 40 or more weeks. 

```{r echo=TRUE}
Wages %>% filter(wks < 40)
Wages %>% filter(wks >= 40) %>% ggplot(aes(exp, lwage)) + geom_jitter() + geom_smooth(method = "lm")
```

Comment on the pattern in your plot—what shape is it? Speculate as to why the pattern looks the way it does.

```{r echo=FALSE}
Wages %>% filter(wks >= 40) %>% select(exp, lwage) %>% cor()
Wages %>% filter(wks >= 40) %>% select(exp, wage) %>% cor()
```


Task 3

```{r}
Wages <-
  Wages %>%
  mutate(wage_levels = cut(
    wage,
    c(0, 750, 1500, max(wage)),
    labels = c("Low", "Mid", "High")
  ))
```


The relationship between wage levels and he sex of the individua:


```{r echo=FALSE}
Wages %>% ggplot(aes(exp, lwage, color = wage_levels)) + geom_point() + facet_wrap(~sex)
```



```{r echo=FALSE}

```






