---
title: "Assignment 2"
subtitle: "STAE04/DABN19: Data Visualization"
author: "Roshan Talimi"
date: "`r Sys.Date()`"
papersize: a4
geometry: margin=4cm
colorlinks: true
output:
  pdf_document:
    number_sections: true
---

```{r include = FALSE}
# https://raw.githubusercontent.com/stat-lu/dataviz/main/template.Rmd"
```

```{r setup, include = FALSE}
# Setup options for R Markdown
knitr::opts_chunk$set(
  echo       = TRUE,    # Do not print code
  warning    = FALSE,    # Suppress warnings
  message    = FALSE,    # Suppress messages
  fig.align  = "center", # Center figures
  fig.width  = 2.7,      # Good standard figure width for single-panel figures
  fig.height = 2.4       # Good standard figure height for single-panel figures
)

library(tidyverse)

# Set a theme for ggplot2
theme_set(theme_light(base_size = 15))

# Set options
options(
  digits = 3, # limit the number of significant digits
  width  = 63 # limit the width of code output
)
```

```{r include = FALSE, echo=FALSE}
# https://archive.ics.uci.edu/ml/datasets/Abalone

url <- paste0(
  "https://archive.ics.uci.edu/",
  "ml/machine-learning-databases/abalone/abalone.data"
)
abalone_raw <- read_csv(url, col_names = FALSE)
```

Abalone data set:

```{r include = FALSE}
abalone_raw %>% sample_n(10) %>% knitr::kable("simple")
```


Task 1

* This data was collected in Tasmania, December of 1995.
* There are 4177 observations and 9 variables in this data set.
* The variables are:

1. Sex (nominal)
2. Length (continuous)
3. Diameter (continuous)
4. Height	(continuous)
5. Whole_weight (continuous)
6. Shucked_weight	(continuous)
7. Viscera_weight (continuous)
8. Shell_weight (continuous)
9. Rings (integer)


Task 2

```{r}
abalone <-
abalone_raw %>% 
  rename(sex = X1,
         length = X2,
         diameter = X3,
         height = X4,
         whole_weight = X5,
  shucked_weight = X6,
  viscera_weight = X7,
	shell_weight = X8,
	rings = X9) %>% 
  mutate(sex = fct_recode(sex),
         rings = as.integer(factor(rings)))
```

```{r}
abalone %>% sample_n(5) %>% knitr::kable("simple")
```

I really don't know how to make this longer :(




