---
title: "Assignment2"
subtitle: "STAE04/DABN19: Data Visualization"
author: "Roshan Talimi"
date: "`r Sys.Date()`"
papersize: a4
geometry: margin=4cm
colorlinks: true
output:
  pdf_document:
    number_sections: true
---

```{r include = FALSE}
# https://raw.githubusercontent.com/stat-lu/dataviz/main/template.Rmd"
```

```{r setup, include = FALSE}
# Setup options for R Markdown
knitr::opts_chunk$set(
  echo       = FALSE,    # Do not print code
  warning    = FALSE,    # Suppress warnings
  message    = FALSE,    # Suppress messages
  fig.align  = "center", # Center figures
  fig.width  = 2.7,      # Good standard figure width for single-panel figures
  fig.height = 2.4       # Good standard figure height for single-panel figures
)

library(tidyverse)

# Set a theme for ggplot2
theme_set(theme_light(base_size = 15))

# Set options
options(
  digits = 3, # limit the number of significant digits
  width  = 63 # limit the width of code output
)
```

```{r include = FALSE}
# https://archive.ics.uci.edu/ml/datasets/Abalone

url <- paste0(
  "https://archive.ics.uci.edu/",
  "ml/machine-learning-databases/abalone/abalone.data"
)
abalone_raw <- read_csv(url, col_names = FALSE)
```


Task 1

* Where and when was this data collected: Tasmania December 1995  
* How many observations and variables are there - 4177 observations, 9 variables  
* What are the variables? Describe each in terms of its units of measurement and it’s data type (continuous, integer, or categorical/nominal).

	Name		Data Type	Meas.	Description
	----		---------	-----	-----------
	Sex		  nominal			M, F, and I (infant)
	Length	continuous	mm	Longest shell measurement
	Diametercontinuous	mm	perpendicular to length
	Height	continuous	mm	with meat in shell
	Whole   weight	continuous	grams	whole abalone
	Shucked weight	continuous	grams	weight of meat
	Viscera weight	continuous	grams	gut weight (after bleeding)
	Shell   weight	continuous	grams	after being dried
	Rings		integer			+1.5 gives the age in years


```{r einclude = FALSE}
colnames(abalone_raw)
```


Task 2

```{r include = FALSE}
abalone_raw

abalone_raw %>% 
  rename(sex = X1,
         length = X2,
         diameter = X3,
         height = X4,
         whole_weight = X5,
  shucked_weight = X6,
  viscera_weight = X7,
	shell_weight = X8,
	rings = X9) %>% 
  mutate(sex = fct_recode(sex),
         rings = as.numeric(factor(rings)))
                       
                  

# abalone <-
#   abalone_raw %>%
#   rename() %>% # rename
#   mutate(NA)     # format
```

Pivot your data into a long format and store this data set in abalone_long using pivot_longer().

```{r}
# abalone_long <-
#   abalone %>%
#   pivot_longer(NA) # pivot your data
```


```{r, fig.cap = "Average weight of Tatooine born Star Wars charachters by eye color."}
# Plot these variables using both box and violin plots.
```


